---
title: 'Project: Music Store'
author: 'Michał Lubczyński'
output:
  pdf_document: default
---


```{r include = FALSE}
library(DBI)
library(knitr)
library(kableExtra)

#Connection
db <- 'dsqu_musicstore' 
host_db <- #University database
db_port <- #Port
db_user <- rstudioapi::askForPassword("Database user")   
db_password <- rstudioapi::askForPassword("Database password")
con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password)
```



**a) Provide a list of artists along with the number of genres they play. List sorted in descending order by musical genre and ascending by artist name. Additionally, in the third column give the place in dense rank in descending order. Limit results to 20 records.**
```{sql connection=con, output.var = "a", echo = TRUE, results = "hide"}
SELECT DISTINCT a.name "Performer_name", 
        count(DISTINCT g.genreid) "Music_genre_count",
        DENSE_RANK() over(ORDER BY count(DISTINCT g.genreid) DESC) "Rank"
FROM artist a
JOIN album al ON a.artistid = al.artistid 
JOIN track t ON al.albumid = t.albumid 
JOIN genre g ON t.genreid = g.genreid
GROUP BY a.name
ORDER BY "Music_genre_count" DESC
LIMIT 20
```
```{r echo = FALSE}
kableExtra::kable(a, booktabs = T, longtable = T) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "hold_position")
```



**b) What kinds of music genres do Canadian customers listen to? Display genre and number of distinct songs in the genre.**
```{sql connection=con, output.var = "b", echo = TRUE, results = "hide"}
SELECT DISTINCT g.name "Genre",
        count(DISTINCT t.trackid) "Song_count"
FROM genre g 
JOIN track t ON g.genreid = t.genreid
JOIN invoiceline i ON t.trackid = i.trackid
JOIN invoice inv ON i.invoiceid = inv.invoiceid
WHERE inv.billingcountry = 'Canada'
GROUP BY "Genre"
ORDER BY "Song_count" DESC
```
```{r echo = FALSE}
kableExtra::kable(b, booktabs = T, longtable = T) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "hold_position")
```



**c) List the ten customers who spent the most in this store.**
```{sql connection=con, output.var = "c", echo = TRUE, results = "hide"}
SELECT  firstname,
        lastname,
        sum(total)
FROM invoice
JOIN customer ON invoice.customerid = customer.customerid
GROUP BY invoice.customerid, firstname, lastname
ORDER BY sum(total) DESC
LIMIT 10
```
```{r echo = FALSE}
kableExtra::kable(c, booktabs = T, longtable = T) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "hold_position")
```



**d)Give the distribution of the sum of money spent with the breakdown by customer countries as a percentage with an accuracy of one per mille. Sort the result by the largest share.**
```{sql connection=con, output.var = "d", echo = TRUE, results = "hide"}
SELECT billingcountry,
        round(sum(total)*100/(SELECT sum(total) FROM invoice), 1) "%_share"
FROM invoice
GROUP BY billingcountry
ORDER BY sum(total) DESC
```
```{r echo = FALSE}
kableExtra::kable(d, booktabs = T, longtable = T) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "hold_position")
```



**e) Display the percentage (with an accuracy of two decimal places) of the types of music file formats purchased from the entire data set (and separately by genre). What genre of music has nobody bought?**
```{sql connection=con, output.var = "e", echo = TRUE, results = "hide"}
WITH cte_counter AS 
        (SELECT gen.name genre_name,
                med.name format_name,
                count(inv.trackid) OVER(PARTITION BY med.name) format_song_count,
                round((count(inv.trackid) OVER 
                        (PARTITION BY med.name)*100.0 / 
                        (SELECT count(trackid) FROM invoiceline)), 2) "%_format",
                count(inv.trackid) over(PARTITION BY gen.name, med.name) genre_song_count,
                round((count(inv.trackid) OVER 
                        (PARTITION BY gen.name, med.name)*100.0 / 
                        (SELECT count(trackid) FROM invoiceline)), 2) "%_genre"
        FROM invoiceline AS inv
        JOIN track t ON inv.trackid = t.trackid
        JOIN mediatype med ON t.mediatypeid = med.mediatypeid
        RIGHT JOIN genre gen ON t.genreid = gen.genreid)
SELECT DISTINCT *
FROM cte_counter
ORDER BY format_song_count DESC , genre_song_count DESC
```
```{r echo = FALSE}
kableExtra::kable(e, booktabs = T, longtable = T) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "hold_position", full_width=T, font_size=9)
```



**f) What artist is most often bought by those who also bought Miles Davis' albums (excluding Miles Davis and Various Artists)**
```{sql connection=con, output.var = "f", echo = TRUE, results = "hide"}
SELECT art.name,
        count(*) "song_count"
FROM artist art
    JOIN album al ON art.artistid = al.artistid
    JOIN track tr ON al.albumid = tr.albumid
    JOIN invoiceline inv ON tr.trackid = inv.trackid
    JOIN invoice i ON inv.invoiceid = i.invoiceid
    JOIN customer cus ON i.customerid = cus.customerid
WHERE cus.customerid IN (SELECT cus.customerid "Miles_fan"
                            FROM artist art
                                JOIN album al ON art.artistid = al.artistid
                                JOIN track tr ON al.albumid = tr.albumid
                                JOIN invoiceline inv ON tr.trackid = inv.trackid
                                JOIN invoice i ON inv.invoiceid = i.invoiceid
                                JOIN customer cus ON i.customerid = cus.customerid
                            WHERE art.name LIKE 'Miles%')
                            AND art.name NOT LIKE 'Miles%'
                            AND art.name NOT LIKE 'Various%'
GROUP BY art.name
ORDER BY count(*) DESC
LIMIT 1
```
```{r echo = FALSE}
kableExtra::kable(f, booktabs = T) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "hold_position")
```

\pagebreak

**g) Create a ranking of customer service representatives (supportrepid), where the evaluation criterion is the obtained highest non-zero turnover for a given month. The query should return a table with the name and surname of the employee of the month for each sales month.**
```{sql connection=con, output.var = "g", echo = TRUE, results = "hide"}
WITH cte AS 
        (SELECT EXTRACT(YEAR FROM invoicedate) "Year",
                EXTRACT(MONTH FROM invoicedate) "Month",
                supportrepid "Employee_id",
                employee.firstname "Employee_firstname",
                employee.lastname "Employee_lastname",
                sum(total) "Sale_sum",
                rank() OVER(PARTITION BY EXTRACT(YEAR FROM invoicedate),
                                            EXTRACT(MONTH FROM invoicedate)
                                            ORDER BY sum(total) DESC) "Emp_Rank"
            FROM invoice 
                JOIN customer ON invoice.customerid = customer.customerid
                JOIN employee ON customer.supportrepid = employee.employeeid
            GROUP BY "Year", "Month", "Employee_id", 
                      "Employee_firstname", "Employee_lastname"
            ORDER BY "Year", "Month", "Sale_sum" DESC)         
SELECT "Year",
        "Month",
        "Employee_id",
        "Employee_firstname",
        "Employee_lastname",
        max("Sale_sum")
FROM cte
WHERE "Emp_Rank" = 1
GROUP BY "Year", "Month", "Employee_id", 
          "Employee_firstname", "Employee_lastname"
ORDER BY "Year", "Month"
```
```{r echo = FALSE}
kableExtra::kable(g, booktabs = T, longtable = T) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "hold_position")
```



**h) For each of the three employees (supportrepid), indicate the months for which they recorded zero sales.**
```{sql connection=con, output.var = "h", echo = TRUE, results = "hide"}
WITH cte AS 
        (SELECT cus.supportrepid "employee",
                EXTRACT (YEAR FROM inv.invoicedate) "year",
                EXTRACT (MONTH FROM inv.invoicedate) "month"
        FROM customer cus
            JOIN invoice inv ON cus.customerid = inv.customerid
        ORDER BY "employee", "year", "month")
SELECT sub1.employee,
        sub1.year,
        CASE WHEN sub1.month = 12 THEN 1
                ELSE sub1.month+1 END AS "month"
FROM (SELECT "employee",
                "year",
                "month",
                lead("month") OVER (ORDER BY "employee") "lead1"
        FROM cte) "sub1"
WHERE sub1.lead1 - sub1.month > 1 OR sub1.lead1 - sub1.month = -10
GROUP BY sub1.employee, sub1.year, sub1.month, sub1.lead1
ORDER BY sub1.employee, sub1.year, sub1.month
```
```{r echo = FALSE}
kableExtra::kable(h, booktabs = T) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "hold_position")
```